---
title: 'Lab 1: Intro to R'
author: "Avery Davidowitz"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Exercise 1

```{r view-girls-counts}
arbuthnot$girls
```

### Exercise 2

Excluding a period of sharp decline centered around 1645 the number of girls baptized shows a steady slow increase as shown in the plot below.

```{r trend-girls}
ggplot(data = arbuthnot, aes(x = year, y = girls)) + 
  geom_line()
```

### Exercise 3

The proportion of boys born over time is consistently higher than girls. Generally it vacillates between .5 and 3.5 percent higher than girls born as shown in the plot below.

```{r plot-prop-boys-arbuthnot}
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)
arbuthnot <- arbuthnot %>%
  mutate(percent_diff = 100 * abs(boys - girls) / total / 2)
ggplot(data = arbuthnot, aes(x = year, y = percent_diff)) + 
  geom_point()
```

### Exercise 4

The data frame is 82 rows by 3 columns: year, boys and girls. The year spans from 1940 to 2002.

```{r dim-present}
glimpse(present)
present %>%
  summarize(min = min(year), max = max(year))
```

### Exercise 5

As demonstrated by the average total number of children counted in a given year in each data frame, the US births were over 300 times higher than the London births.

```{r count-compare}
present <- present %>%
  mutate(total = boys + girls)
present %>%
  summarize(average = mean(total))
arbuthnot %>%
  summarize(average = mean(total))
```

### Exercise 6

Similar to the Arbuthnot data frame, boys are consistently born more commonly. However, in this data the difference in representation is noticeably less spread and stays between a 1% and 1.5% difference.

```{r plot-prop-boys-present}
present <- present %>%
  mutate(total = boys + girls)
present <- present %>%
  mutate(percent_diff = 100 * abs(boys - girls) / total / 2)
ggplot(data = present, aes(x = year, y = percent_diff)) + 
  geom_point()
```

### Exercise 7

1961 saw the most total births in the US between 1940 and 2002.

```{r find-max-total}
present %>%
  arrange(desc(total))
```
