---
title: DATA 606 Data Project Proposal
author: Avery Davidowitz
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data

```


### Research question 

How influential on current Covid deaths are historical vaccinations?



### Cases 

There are 239 cases.



### Data collection 

Local health organizations from countries around the world report Covid cases, deaths and vaccination data to the World Health Organization.



### Type of study 

The data is observational in nature.



### Data Source 

The data can be downloaded from the WHO at https://covid19.who.int/data




### Dependent Variable

"Deaths - newly reported in last 7 days" is a quantitative dependent variable.



### Independent Variable(s)

The independent variables are:
TOTAL_VACCINATIONS_PER100,	PERSONS_VACCINATED_1PLUS_DOSE_PER100,	PERSONS_FULLY_VACCINATED_PER100,	FIRST_VACCINE_DATE,	NUMBER_VACCINES_TYPES_USED, PERSONS_BOOSTER_ADD_DOSE_PER100



### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}

```