---
title: "DATA 606 Data Project Proposal"
author: "Avery Davidowitz"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Import Libraries

```{r}
library(tidyverse)

```


### Data Preparation

```{r}
# load data
url <- "https://raw.githubusercontent.com/adavidowitz100/DATA606/main/FinalProject/email_data.csv"

raw_email_data <- readr::read_csv(url, col_types = list(
              Seniority = col_factor(),
              Function = col_factor(),
              "Subject Line" = col_character(),
              "Hook 1" = col_character(),
              "Z_Contact Status" = col_factor(),
              "Email Opened" = col_date(format = "%m/%d/%Y"),
              "Email Replied" = col_date(format = "%m/%d/%Y")
))

df <- raw_email_data |> dplyr::rename(seniority = Seniority,
                                      department = Function, 
                                      subject = "Subject Line",
                                      hook = "Hook 1", 
                                      status = "Z_Contact Status",
                                      opened_date = "Email Opened",
                                      replied_date = "Email Replied") |>
                      dplyr::mutate(opened = if_else(!is.na(opened_date), TRUE, FALSE)) |>
                      dplyr::mutate(replied = if_else(!is.na(replied_date), TRUE, FALSE)) |>
                      dplyr::mutate(days_diff = difftime(opened_date, replied_date, units = "days"))
head(df)

```


### Research question 

Does the recipient's seniority and department influence the effectiveness of this email marketing campaign? If the current campaign in not equally effective for all recipients the next campaign could be designed to be targeted more effectively by seniority and department.


### Cases 

There are 1449 cases/observations representing marketing emails sent to specific people.


### Data collection 

Real sanitized email market research data was used.


### Type of study 

The data is observational in nature.


### Data Source 

The data is from a business to business email market research campaign. The data can be viewed on Github at https://raw.githubusercontent.com/adavidowitz100/DATA606/main/FinalProject/email_data.csv


### Dependent Variable

The dependent variables are "opened", "replied" and "days_diff" representing if an email was opened, responded to and the days between those events. 


### Independent Variable(s)

The independent variables are department and seniority.


### Relevant summary statistics 

```{r}
seniority_open_xtab <- xtabs(~ seniority + opened, data=df)
seniority_open_xtab
prop.table(seniority_open_xtab, 1)
seniority_replied_xtab <- xtabs(~ seniority + replied, data=df)
seniority_replied_xtab
prop.table(seniority_replied_xtab, 1)

ggplot(df, aes(x=days_diff)) + geom_histogram()
```