---
title: "Data 606 Lab 6 - Inference for categorical data"
author: "Avery Davidowitz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Load Packages


```{r cars}
library(tidyverse)
library(openintro)
library(infer)
```

Data set
```{r}
data('yrbss', package='openintro')
head(yrbss)
```

## Exercise 1
The counts within each category for the amount of days these students have texted while driving within the past 30 days.
```{r}
texting_counts <- yrbss |> dplyr::filter(!is.na(text_while_driving_30d)) |> count(text_while_driving_30d)

texting_counts
```

## Exercise 2
The proportion of people who have texted while driving every day in the past 30 days and never wear helmets.

```{r}
never_helm_always_text <- yrbss |> dplyr::filter(text_while_driving_30d == "30" & helmet_12m == "never") |> nrow()
never_helm_always_text_prop <- never_helm_always_text / nrow(yrbss)
never_helm_always_text_prop
```


```{r}
no_helmet <- yrbss %>%
  filter(helmet_12m == "never")
no_helmet <- no_helmet %>%
  mutate(text_ind = ifelse(text_while_driving_30d == "30", "yes", "no"))
no_helmet <- no_helmet |> filter(!is.na(text_ind)) # I was getting errors without removing NAs
no_helmet %>%
  specify(response = text_ind, success = "yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)
```

## Exercise 3
The margin of error for the estimate of the proportion of non-helmet wearers that have texted while driving each day for the past 30 days based on this survey.
```{r}
margin_error <- 1.96 * (never_helm_always_text_prop * (1 - never_helm_always_text_prop) / nrow(yrbss))^(1/2)
margin_error
```

## Exercise 4
Calculated confidence intervals for two other categorical variables.
```{r}
hispanic_pop_ci <- yrbss |> filter(!is.na(hispanic)) %>%
  specify(response = hispanic, success = "hispanic") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)
hispanic_pop_ci
```
I can claim with 95% confidence that the true Hispanic proportion of the general population of teens is between 24.9% and 26.4%. 
```{r}
gender_ci <- yrbss |> filter(!is.na(gender)) |>
  specify(response = gender, success = "female") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)
gender_ci
```
I can claim with 95% confidence that the true female proportion of the general population of teens is between 47.9% and 49.6%.

## Exercise 5

The margin of error is the largest when p = .5 and the smallest when p approaches 0 and 1. However, when p is very small or very large it may be difficult to obtain a large enough sample to satisfy the success-failure conditions needed to use a normal distribution.

```{r}
n <- 1000
p <- seq(from = 0, to = 1, by = 0.01)
me <- 2 * sqrt(p * (1 - p)/n)
dd <- data.frame(p = p, me = me)
ggplot(data = dd, aes(x = p, y = me)) + 
  geom_line() +
  labs(x = "Population Proportion", y = "Margin of Error")
```


```{r sf-app, echo=FALSE, eval=FALSE}
library(shiny)
shinyApp(
  ui = fluidPage(
      numericInput("n", label = "Sample size:", value = 300),
      
      sliderInput("p", label = "Population proportion:",
                  min = 0, max = 1, value = 0.1, step = 0.01),
      
      numericInput("x_min", label = "Min for x-axis:", value = 0, min = 0, max = 1),
      numericInput("x_max", label = "Max for x-axis:", value = 1, min = 0, max = 1),
    plotOutput('plotOutput')
  ),
  
  server = function(input, output) { 
    output$plotOutput = renderPlot({
      pp <- data.frame(p_hat = rep(0, 5000))
      for(i in 1:5000){
        samp <- sample(c(TRUE, FALSE), input$n, replace = TRUE, 
                       prob = c(input$p, 1 - input$p))
        pp$p_hat[i] <- sum(samp == TRUE) / input$n
      }
      bw <- diff(range(pp$p_hat)) / 30
      ggplot(data = pp, aes(x = p_hat)) +
        geom_histogram(binwidth = bw) +
        xlim(input$x_min, input$x_max) +
        ggtitle(paste0("Distribution of p_hats, drawn from p = ", input$p, ", n = ", input$n))
    })
  },
  
  options = list(height = 500)
)
```

## Exercise 6
p hat is narrowly distributed around .1 with a nearly normal shape.

## Exercise 7
The distribution of p hats seem very similar for most ranges of p. p hat's distribution does seem to approximate the normal distribution the closer p is to .5. Many other ranges of values of p generate multi-modal distributions.
## Exercise 8
The higher the sample size the tighter the spread and the more normal shape for p hat. The distribution of p hat seems more sensitive to p than to sample size.

## Exercise 9
Is there convincing evidence that those who sleep 10+ hours per day are more likely to strength train every day of the week?
I would test a H naught of p1 - p0 = 0. Where p1 is the proportion of those who sleep 10+ and strength train everyday / pop and p0 is the proportion of those who strength train everyday but do not sleep 10+ / pop.
```{r}
pop_size <- nrow(yrbss)
sleep_over10_strevery <- yrbss |> dplyr::filter(school_night_hours_sleep == "10+") |>
                        dplyr::count(strength_training_7d)
sleep_over10_size <- yrbss |> dplyr::filter(school_night_hours_sleep == "10+") |> nrow()
yrbss |> dplyr::count(strength_training_7d)
```

## Exercise 10
Alpha represents the probability that you reject the null hypothesis incorrectly or a type 1 error. Therefore there would be a 5% chance that your statistical analysis could falsely determine a difference in the likelihood to strength train everyday given sleeping over 10+ hours.     

## Exercise 11
You can establish an upper bound on the margin of error if you solve for n using a p = q = .5 which produces the highest margin of error. ME = Z × SE -> n = :

```{r}
n <- (.5*1.96/.01)^2
n
```

